Almost orthogonal vectors

### 2014-12-10 project3/paper/paper.tex
Paper changes

Along the same lines, given a configuration of $N$ vectors in $n$-space with corresponding $\epsilon$, we can obtain a collection of $N+1$ vectors in $\mathbb R^{n+1}$ with the same $\epsilon$: simply embed the first $N$ vectors in $\mathbb R^{n}$, then add any vector perpendicular to the hyperplane in which the $N$ vectors lie. Hence the minimum $\epsilon_3$ for collections of $N+1$ vectors in $\mathbb R^{n+1}$ is upper-bounded by the corresponding $\epsilon$ for $N$ vectors in $\mathbb R^n$.

@@

Observe that as before, we are interested in maximizing the square of the cosine of the angle between any two vectors in $v_1, v_2, \ldots, v_N$. Note that here, the angle between any two vectors in $v_j$ and $v_k$ in $v_1, v_2, \ldots, v_N$ can be expressed as $\cos^2 (\sum_{i=j} ^{k-1} \theta_i)$. Note that, however, if $\sum_{i=j}^{k-1} \theta_i \leq \pi/2$, then $\cos^2 \theta_j \geq \cos^2(\sum_{i=j}^{k-1} \theta_i)$ and that if $\sum_{i=j}^{k-1} \theta_i > \pi/2$, then $\cos^2 (\sum_{i=j}^{k-1} \theta_i) = \cos^2 (\pi - \sum_{i=j}^{k-1} \theta_i)$ which is less than $\cos^2 \theta_i$ for any $i$ not in $\{j, j+1, \ldots, k-1\}$.

@@

Figure $3$ shows an optimum configuration for $n=2$ and $N=5$. Note by taking reflections of $x_2$ and $x_4$ in Figure $3$, we get five vectors that form a regular $5$-gon -- this can be seen in Figure $4$. A similar transformation can be done for all odd $N$; unfortunately, even $N$ do not produce regular polygons.

@@

\section{Simulated annealing and associated conjectures}

Since visualizing in spaces of dimensionality greater than $3$ is extremely hard, we used an annealing approach to compute near-optimal epsilon values for different $(n, N)$ pairs. We describe this method in greater detail below.

\begin{lstlisting}
def get_min_epsilon_annealing(n, N, num_iter, start_vectors):
    min_epsilon = 1.0  # Start off with the worst possible epsilon
    current_vectors = start_vectors
    i = 0
    while (i < num_iter):
        temperature = 1.0 / float(i + 1)

@@

    return min_epsilon, current_vectors
\end{lstlisting}
\captionof{lstlisting}{The annealing algorithm used to generate optimum epsilon values}

@@

Even though this approach seems simple, it actually produces remarkably precise values -- this can be verified by cross-checking with results we've already proven / conjectured in this paper.

To make this more concrete, consider Table $1$.

The accuracy of our method can be noted by looking at the computed values for $n=2$ -- the computed values for different values of $N$ are very close to the expected value of $\cos^2 \pi/N$, that we already proved in Section $2$, which makes us confident about the accuracy of this numeric method.

>>>>

Along the same lines, given a configuration of $N$ vectors in $n$-space with corresponding $\epsilon$, we can obtain a collection of $N+1$ vectors in $\mathbb R^{n+1}$ with the same $\epsilon$: simply embed the first $N$ vectors in $\mathbb R^{n}$, then add any vector perpendicular to the hyperplane in which the $N$ vectors lie. Hence the minimum $\epsilon$ for collections of $N+1$ vectors in $\mathbb R^{n+1}$ is upper-bounded by the corresponding $\epsilon$ for $N$ vectors in $\mathbb R^n$.

@@

Observe that as before, we are interested in maximizing the square of the cosine of the angle between any two vectors in $v_1, v_2, \ldots, v_N$. Note that here, the angle between any two vectors in $v_j$ and $v_k$ in $v_1, v_2, \ldots, v_N$ can be expressed as $\sum_{i=j} ^{k-1} \theta_i$. Note that, however, if $\sum_{i=j}^{k-1} \theta_i \leq \pi/2$, then $\cos^2 \theta_j \geq \cos^2(\sum_{i=j}^{k-1} \theta_i)$ and that if $\sum_{i=j}^{k-1} \theta_i > \pi/2$, then $\cos^2 (\sum_{i=j}^{k-1} \theta_i) = \cos^2 (\pi - \sum_{i=j}^{k-1} \theta_i)$ which is less than $\cos^2 \theta_i$ for any $i$ not in $\{j, j+1, \ldots, k-1\}$.

@@

Figure $3$ shows an optimum configuration for $n=2$ and $N=5$. Note by taking reflections of $x_2$ and $x_4$ in Figure $3$, we get five vectors that form a regular $5$-gon -- this can be seen in Figure $4$. A similar transformation can be done for all odd $N$.

@@

\section{Numerical analysis and associated conjectures}

Since visualizing in spaces of dimensionality greater than $3$ is extremely hard, we used a computational approach to compute near-optimal epsilon values for different $(n, N)$ pairs. We describe this method in greater detail below.

\begin{lstlisting}
def get_min_epsilon(n, N, num_iter, start_vectors):
    min_epsilon = 1.0  # Start off with the worst possible epsilon
    current_vectors = start_vectors
    i = 0
    while (i < num_iter):
        temperature = 1.0 / float(i + 1)

@@

    return min_epsilon, current_vectors
\end{lstlisting}
\captionof{lstlisting}{The numerical method used to generate optimum epsilon values -- note that we call this method multiple times with different start\_vectors to obtain a reasonable epsilon estimate}

@@

Even though this approach seems simple, it actually produces remarkably precise values -- this can be verified by cross-checking with results we've already proven / conjectured in this paper, this can be seen from values in Table $2$.

In particular, the computed values for $n=2$ are pretty accurate -- the computed values for different values of $N$ are very close to the expected value of $\cos^2 \pi/N$, that we already proved in Section $2$, which makes us confident about the accuracy of this numeric method.
